---
title: "Juvenile Snowshoe Hares in Bonanza Creek: Exploratory Report"
author: "Alia Ajina"
date: "11/3/2021"
output: html_document
---

## 1. Introduction

This exploratory report provides a synopsis on juvenile snowshoe hares (*Lepus americanus*) observed at three different sites at the Bonanza Creek Experimental Forest (Long Term Ecological Research site) which is located 20 km southwest of Fairbanks, Alaska. According to the data, snowshoe hares are a 'keystone' prey species in boreal forests and "experience population fluctuations of 8 to 11 years," (Kielland et al. 2017). Using data from 1998 to 2012 that was collected by Kielland et al., this report analyses annual juvenile hare count and juvenile hare weights. It also compares juvenile hare weights for both males and females and performs a linear regression on the relationship between juvenile weight and hind foot length. 

## 2. Data and Analyses

Snowshoe hare physical data was collected and made accessible by Dr. Knut Kielland and his primary investigators and associates at the Bonanza Creek Long Term Ecological Research (LTER) site. The data contains observations from capture-recapture studies that were conducted in the Tanana valley from 1998 - 2012, (Kielland et al. 2017). Following exploratory data visualization, body mass measurements between male and female juvenile snowshoe hares are compared using a two-sample t-test using a significance level (α) of 0.05 throughout. Differences between groups are described by a Cohen's d effect size. The relationship between body mass and hind foot length is explored by simple linear regression. All analyses are in R version 4.1.1 using RStudio Version 1.4.1717.

```{r setup, include = FALSE, message = FALSE, echo = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

# Setup: Attach necessary packages
library(tidyverse)
library(here)
library(janitor)
library(lubridate)
library(kableExtra)
library(effsize)
library(broom)
```

## 3. Juvenile Hare Initial Investigation

```{r reading data}
# Read in Bonanza Creek Hares data
bonanza <- read_csv(here("data", "bonanza_hares.csv"))
```

### A. Annual Juvenile Hare Trap Counts

```{r counts}
# Data wrangling: Count total number of juvenile hare trappings each year in the study
juvenile_hare <- bonanza %>% 
filter(age == "j") %>% 
mutate(date = mdy(date)) %>% 
mutate(year = year(date)) %>% 
group_by(year) %>% 
summarise(hare_count = n())
  
```

```{r summary stats}
# Create table of summary statistics in order to analyse major takeaways of data
juvenile_hare_summary <- juvenile_hare %>% 
  summarise(mean_count = mean(hare_count, na.rm = TRUE), 
            median_count = median(hare_count, na.rm = TRUE), 
            sd_count = sd(hare_count, na.rm = TRUE), 
            max_count = max(hare_count, na.rm = TRUE),
            min_count = min(hare_count, na.rm = TRUE))
```

```{r visualiation column}
# Create a finalized data visualization with a column bar graph of counts by year
ggplot(data = juvenile_hare, 
       aes(x = year, y = hare_count)) +
# Change aesthetics accordingly
  geom_col(color = "#964B00", fill = "#964B00") +
  theme_minimal() +   
  scale_x_continuous(breaks = 1998:2012) +
  scale_y_continuous(breaks = c(0, 20, 40, 60, 80, 100, 120, 140)) +
  labs( x = "Year",
        y = "Number of Juvenile Hare Traps") +
# Adjust positioning
  theme(axis.text.x = element_text(size = 8, hjust = 0.5),
        axis.text.y = element_text(size = 8),
        axis.title.x = element_text(hjust = 0.5, 
                                    vjust = 0.5),
        plot.caption = element_text(hjust = .9), 
        legend.position = "none")
```

**Figure 1:** Total number of juvenile snowshoe hare trappings per year from 1998 to 2012 at Bonanza Creek Long Term Ecological Research site. Data: Kielland et al. (2017)

The minimum count for juvenile hare trappings is `r juvenile_hare_summary$min_count` and the maximum is `r juvenile_hare_summary$max_count`. The mean annual number of juvenile hares trapped is `r juvenile_hare_summary$mean_count` and the median is `r juvenile_hare_summary$median_count`. In 1999, juvenile snowshoe hare densities were highest. It appears that since then (1999), the number of juvenile hares that have been trapped has decreased, although the trend is not profound. In years 2005 and 2008 a small increase can be seen in the number of juvenile hares trapped in comparison to years prior. There were also no juvenile hare observations in 1998, 2002, and 2009. These counts, although absolute, are impacted by effort (number of days and number of traps in the study each year). Efforts for each year may have been uniform, or they may have varied greatly, which could have affected the data collected between 1998 to 2012. It would be beneficial to the study to quantify the number of observations recorded each year and standardize that number in future analyses in order to have accurate, standardized data, that may better represent the true population of juvenile hares.

### B. Juvenile Hare Weights

```{r data wrangling hare weights}
# Data Wrangling : Create data frame and wrangle data accordingly
hare_weights_sex <- bonanza %>% 
  filter(age == "j") %>% 
  mutate(full_grid = case_when(
    grid == "bonmat" ~ "Bonanza Mature",
    grid == "bonrip" ~ "Bonanza Riparian",
    grid == "bonbs" ~ "Bonanza Black Spruce")) %>% 
  mutate(full_sex = case_when(
    sex == "f" ~ "female",
    sex == "m" ~ "male",
    TRUE ~ "NA")) %>% 
  group_by(weight) %>% 
  summarize(full_sex, full_grid)
```


```{r histogram}
# Data Visualization with histogram
ggplot(data = hare_weights_sex, aes(x = weight, fill = full_sex)) +
  geom_histogram(color = "white", bins = 15) +
  scale_fill_manual(values = c("pink", "#1261A0", "black")) +
  facet_wrap(~ full_grid) +
  theme_grey() +
  labs(x = "Weight (g)",
       y = "Percent of Sample",
       fill = "Sex") +
# Adjust positioning
  theme(axis.text.x = element_text(size = 6, hjust = 0.5),
        axis.text.y = element_text(size = 6),
        axis.title.x = element_text(hjust = 0.5, 
                                    vjust = 0.5),
        plot.caption = element_text(hjust = .8),
        legend.background = element_rect(fill = "light grey"),
        legend.title.align = 0.5)
```

**Figure 2:** Distribution of juvenile hare weights in grams by the percent of the population recorded faceted at each of the three sites (Bonanza Black Spruce, Bonanza Mature, and Bonanza Riparian) at the Bonanza Creek Long Term Ecological Research site. Data: Kielland et al. (2017)

```{r jitter}
# Data visualization with jitter plot
ggplot(data = hare_weights_sex, aes(x = full_sex, y = weight)) +
  geom_jitter(aes(color = full_sex, 
                     fill = full_sex), bins = 10) +
  scale_color_manual(values = c("pink", "#1261A0", "black")) +
  facet_wrap(~ full_grid) +
  theme_grey() +
  labs(x = "Sex",
       y = "Weight (g)") +
# Adjust positioning
  theme(axis.text.x = element_text(size = 6, hjust = 0.5),
        axis.text.y = element_text(size = 6),
        axis.title.x = element_text(hjust = 0.5, 
                                    vjust = 0.5),
        plot.caption = element_text(hjust = .8), 
        legend.position = "none")
```

**Figure 3:** Jitter plot of the dispersion of juvenile hare weights (g) for each sex faceted by the survey site at Bonanza Creek Long Term Ecological Research site. Data: Kielland et al. (2017)

On average at each of the sites; Bonanza Black Spruce, Bonanza Mature, and Bonanza Riparian, male and female juvenile hares have similar weight distributions. The Bonanza Riparian site recorded more observations in comparison to the other sites. The distributions of weights for female hares are slightly more normal than the distribution of weights for male hares when analyzing the histogram and jitter plot seen above. However, further analysis is needed to determine if this is the case. The NA juvenile hares that are present at Bonanza Mature and Bonanza Riparian may have not been categorized by their sex due to error in data collection efforts.

### C. Juvenile Weight Comparison for Male and Female Snowshoe Hares

```{r wrangling for comparison}
# Data wrangling
sex_weights_sum <- bonanza %>% 
  filter(age == "j") %>% 
  mutate(full_sex = case_when(
    sex == "f" ~ "Female",
    sex == "m" ~ "Male")) %>% 
  group_by(full_sex) %>% 
  summarise(weight)
  
# Create summary statistics
sex_weights_summary <- sex_weights_sum %>% 
  summarize(mean_weight =round(mean(weight, na.rm =TRUE), 2),
            median_weight = median(weight, na.rm = TRUE),
            sd_weight = round(sd(weight, na.rm = TRUE), 2),
            sample_size_weight = n())

# Finalized table:
sex_weights_summary %>% 
  kable(col.names = c( "Sex",
                      "Mean Weight (g)",
                      "Median Weight (g)",
                      "Standard Deviation (g)",
                      "Sample Size"),
        caption = "<center>Table 1: Summary statistics of male and female juvenile snowshoe hare's weight (g) at Bonanza Creek LTER<center>") %>% 
  kable_styling(bootstrap_options = "bordered",
                full_width = F, position = "center")

```

```{r exploratory graphs, include = FALSE}
# Exploratory graphs to decide if using a t-test is appropriate

# Exploratory histogram
ggplot(data = sex_weights_sum, aes(x = weight)) +
  geom_histogram(bins = 10) +
  facet_wrap(~ full_sex)

# Exploratory QQ-plot
ggplot(data = sex_weights_sum, aes(sample = weight)) +
  stat_qq() + 
  stat_qq_line() +
  facet_wrap(~ full_sex)
```

```{r vector creation}
# Means comparison

# Create female vector for means comparison
fem_weight <- sex_weights_sum %>% 
  filter(full_sex == "Female") %>% 
  drop_na(weight) %>% 
  pull(weight)

# Create male vector for means comparison
m_weight <- sex_weights_sum %>% 
  filter(full_sex == "Male") %>% 
  drop_na(weight) %>% 
  pull(weight)
```

```{r cohens d, include = FALSE}
# Means comparison continued

# Run cohen's d for effect size
weight_effect <- cohen.d(fem_weight, m_weight)
# View effect size
weight_effect
```

```{r ttest}
# Means comparison continued

# Run a t-test
weight_t <- t.test(fem_weight, m_weight)
# Get the tidy model results to call outputs in-line:
weight_t_tidy <- tidy(weight_t)
```

```{r actual difference in means}
# Run code to get the means and SDs for both female and male vectors in order to calculate the actual difference in means

fem_mean <- mean(fem_weight)
m_mean <- mean(m_weight)
fem_sd <- sd(fem_weight)
m_sd <- sd(m_weight)
```

Here we compared the juvenile weights of male and female snowshoe hares using the actual difference in means, a Cohen's d test to identify the effect size, and an unpaired two-sample t-test.

On average, male juvenile hares are bigger in weight than females with the mean for males being `r round(m_mean,2)` $\pm$ `r round(m_sd,2)` and mean for females being `r round(fem_mean,2)` $\pm$ `r round(fem_sd,2)`. The absolute difference in means is `r round(m_mean, 2) - round(fem_mean, 2)` mm (a `r round((m_mean - fem_mean)/((m_mean + fem_mean)/2)*100, 2)`% difference). The difference in means is significant (Welch's two-sample t-test: t(`r round(weight_t_tidy$parameter,2)`) = `r round(weight_t_tidy$statistic,2)`, p < 0.01), and the effect size is small (Cohen's *d* = `r round(weight_effect$estimate,2)`)

Getting a p-value of `r round(weight_t[["p.value"]], 3)` means there is a `r round (weight_t[["p.value"]] * 100, 1)`% chance of getting means at least as different as the ones seen here if they're drawn from populations with the same mean. This means that we would reject the null hypothesis that the population's mean weights are the same.
 
 
### D. Relationship between Juvenile Weight & Hind Foot Length

```{r data wrangling weight and hind foot length}
# Wrangle data to analyze relationship between juvenile weight and hind foot length
hare_weight_hindft <- bonanza %>%
filter(age == "j") %>% 
  select(hindft, weight)
```

```{r visualization scatter plot}
# Create a final visualization of juvenile hare hind foot length vs weight
ggplot(data = hare_weight_hindft, aes(x = hindft, y = weight)) +
  geom_point(color = "#964B00", size = 2) +
  geom_smooth(method = "lm", color = "black") +
  theme_minimal() +
  labs(x = "Hind Foot Length (mm)", 
       y = "Weight (g)" ) + 
# Adjust positioning
  theme(axis.text.x = element_text(size = 6, hjust = 0.5),
        axis.text.y = element_text(size = 6),
        axis.title.x = element_text(hjust = 0.5, 
                                    vjust = 0.5),
        plot.caption = element_text(hjust = .8))
```

**Figure 4:** Visual relationship between all juvenile snowshoe hare's hind foot length (mm) and weight (g).


```{r linear regression, include = FALSE}
# Perform a linear regression to describe how juvenile snowshoe hare weight changes with respect to hind foot length
lm_hare <- lm(weight ~ hindft, data = hare_weight_hindft)

# Use tidy to get a tidy output of lm
lm_hare_tidy <- tidy(lm_hare)
# Display tidy output of lm
lm_hare_tidy

# Get coefficient
hare_coef <- lm_hare_tidy$estimate[2]

# Use glance to get output of model fit & statistical significance
lm_hare_glance <- glance(lm_hare)
# Display output of model fit & statistical significance 
lm_hare_glance
```

```{r exploratory vis, include = FALSE}
# Exploratory visualizations
plot(lm_hare)
```

```{r pearsons r, inculde = FALSE}
# Pearson's r
hare_cor <- cor.test(hare_weight_hindft$weight, hare_weight_hindft$hindft)

# View result
invisible(hare_cor)
```

Linear regression exploration was used to further describe the relationship between juvenile snowshoe hare's weight changes (g) with respect to hind foot length (mm) for all juvenile hares within the data set. A regression model was found ($\beta$ = `r round(hare_coef,3)`, F(`r lm_hare_glance$df`,`r lm_hare_glance$df.residual`) = `r round(lm_hare_glance$statistic,1)`, p < 0.001) with an R^2^ of `r round(lm_hare_glance$r.squared,3)`.

With this linear regression, the R^2^ value of `r (lm_hare_glance$r.squared)` means that only `r (lm_hare_glance$r.squared) * 100`% of the variance in body mass (g) is predictable from the change in hind foot length (mm). 

Using a Pearson's R correlation, it is shown that there is a *moderately strong* correlation between juvenile hare hind foot length (mm) and body mass (g) (*r* = `r round(hare_cor$estimate,2)`, t(`r hare_cor$parameter`) = `r round(hare_cor$statistic,2)`, p < 0.001). 

Certain variables, for example, the sex of juvenile hares, is not addressed in this linear regression, which may be cause for concern as it may impact the relationship between hare weight changes and hind foot length.

## Summary: 

- The first year of the study, 1999, contained the highest density of juvenile hares trapped.
- On average, female juvenile snowshoe hares are smaller than male juvenile snowshoe hares in weight. Male juvenile snowshoe hares have a mean of `r round(m_mean, 2)` and female juvenile snowhoe hares have a mean of `r round(fem_mean, 2)`.
- Juvenile hare weight seems to be related to hind foot length. The simple linear regression between the two has a slope of `r round(lm_hare_tidy$statistic[2], 2)`. In other words, for every 1mm increase in hind foot length, on average, it is expected that weight will increase by `r round(lm_hare_tidy$statistic[2], 2)`.

## Next Steps:

- Investigating the normality of female juvenile hare weights versus male juvenile hare weights distributions
- Adding potential predictor variables to the linear regression such as sex of juvenile hares
- Investigating all analyses for adult snowshoe hares to see if percieved relationships continue

## Citations: 

Kielland, K., F.S. Chapin, R.W. Ruess, and Bonanza Creek LTER. 2017. Snowshoe hare physical data in Bonanza Creek Experimental Forest: 1999-Present ver 22. Environmental Data Initiative. https://doi.org/10.6073/pasta/03dce4856d79b91557d8e6ce2cbcdc14

